---
title: "Building Web Applications with Shiny in R"
subtitle: "DataCamp Course"
output: 
  html_notebook:
    number_sections: TRUE
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval=FALSE)
```


# Getting Started with Shiny

- What is a web app?
  - Uses a UI and a server
  
- Most people use `fluidPage()` to generate the UI
- Server created by defining custom function
- Basic structure below

```{r}
library(shiny)

ui <- fluidPage()

server <- function(input, output, session){
  
}

shinyApp(ui = ui, server = server)
```

- To start to add variables the user can input
  - Have to assign inputs to output objects
  - Need to tell the UI to display object calculated by server

```{r}
ui <- fluidPage(
  titlePanel("Title"),
  textInput(
    inputId = "name",
    label = "Enter a name:"
  ),
  textOutput("q")
)

server <- function(input, output, session) {
  output$q <- renderText({
    paste(
      "Do you prefer dogs or cats,",
      input$name,
      "?"
    )
  })
}
```

- To change the layout of the UI, need to add sidebar panels and sidebar layout
- Can use subsets of the dataframe to create plots that change based on the input values

```{r}
ui <- fluidPage(
  titlePanel("Title"),

  sidebarLayout(
    
    sidebarPanel(
      textInput(
        inputId = "name",
        label = "Enter a name:"
      )
    ),

    mainPanel(
      plotOutput("trend")
    )
  )
)

server <- function(input, output, session) {
  output$trend <- renderPlot({
    data_name <- filter(babynames, name == input$name)
    
    ggplot(data_name, aes(x = year, y = prop, color = sex)) +
      geom_line()
  })
}
```

# Inputs, Outputs, and Layouts

